---
title: "Standard Populations"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{standard_populations}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE, 
  message = FALSE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)
```

# Introduction
In this vignette, we will explore the EpiStandard functions designed to retrieve or create a standard population data table. There are two functions to help with this:

- `standardPopulation()`
- `ageGroups()`

## Retrieve standard population data

The package contains two standard population data tables:

- ESP2013: 
- WSP2025:
```{r}
library(EpiStandard)
library(gt)

pop <- standardPopulation("esp2013")
gt(pop)
```

You can create new age groups using the 'mergeAgeGroups()' function. This allows you to combine the age groups in the standard population.

```{r merge}
merged_data <- mergeAgeGroups(pop, c("0-19", "20-64", "65-150"))
gt(merged_data)
```
