---
title: "Adjusting Age Groups"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{standardPopulation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r base, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE, 
  message = FALSE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)
```


# Introduction

This vignette illustrates how `EpiStandard` can be used to adjust age groups in the standard population and in study results. 

# Standard Population

`EpiStandard` includes two standard populations, which are:
- European Standard Population 2013
- World Standard Population 2025

Both use the same age groups, and can be used by using the function `standardPopulation`. You can choose which of these to use by setting the argument `region` to 'Europe' or 'World'.

```{r age groups}
library(EpiStandard)
library(dplyr)

ageGroups <- standardPopulation(region = "Europe") 

ageGroups |>
  pull(age_group)
```

However, some studies might use different age groups which are not compatible with the standard populations. This can be solved by using the function `mergeAgeGroups()`.For example, if a study only uses the age groups '0-19', '20-64' and '65 to 150', the standard population can be adjusted to match

```{r merge groups 1}
newAgeGroups <- mergeAgeGroups(refdata = ageGroups, newGroups = c("0-19", "20-64", "65-150"))

newAgeGroups
```

This will also work if using a bespoke standard population.

```{r merge groups 2}
df_study <- data.frame(age=c('0-14','15-24','25-44','45-64','65-150'),
                       pop=c(114350,80259,133440,142670,92168))

new_df_study <- mergeAgeGroups(refdata = df_study, newGroups = c("0-24", "25-64", "65-150"),
                               age = "age",
                               pop = "pop")

new_df_study |> dplyr::glimpse()
```
# Study Results

Additionally, you can adjust your study results to merge age groups, while taking into consideration additional stratifications of interest. For example, the data set below shows study results for the UK and France. If we want merge some age groups, but still look at each country separately, we can use the argument `strata`. 

```{r merge groups 3}
df_study <- data.frame(country=rep(c('UK',"France"), c(5,5)),
                       age=rep(c('0-14','15-24','25-44','45-64','65-150'),2),
                       deaths=c(132,87,413,2316,3425,605,279,3254,9001,8182),
                       fu=c(114350,80259,133440,142670,92168,37164,20036,32693,14947,2077))

new_df_study <- mergeAgeGroups(refdata = df_study, newGroups = c("0-24", "25-64", "65-150"),
                               age = "age",
                               pop = "fu",
                               event = "deaths",
                               strata = "country")

new_df_study |> dplyr::glimpse()
```
Note: All data used in this vignette is artifical. 
